An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-09-29 paper.tex
Update §5.2.3

While these top participants are influential within the \emph{mailing
lists}, it is unclear how this influence translates into document
authorship and into gaining  leadership roles within  the organisation.

%..................................................................................................
\pb{Draft authorship}
The output of the IETF is technical standards and other documents published
in the RFC series. RFCs are developed by WGs from a sequence of Internet-Drafts,
with individuals acting as named authors.  Having identified the most
influential participants in the mailing list community, we can determine
whether or not these same participants are also the most active authors.

Figure~\ref{fig:drafts_initiation_top_percentile} shows the distribution of
the proportion of documents authored by mailing list participants, sorted
by their influence rank. We observe a growing proportion of the community
involved in draft authoring, skewed towards influential participants who
tend to write more drafts.  During 2010-2019, each participant in the top
10 percentile, authored 0.125\% to 0.175\% of the total number of drafts in
that period. We find that 32\%-42\% of the total drafts, in different years
during this period, were authored by the top 10\% most influential mailing
list participants.

@@

%..................................................................................................
\pb{Co-authorship and email graph correlation}
Figure~\ref{fig:drafts_initiation_top_percentile} showed that influential
mailing list participants tend to author more drafts than others.
We next ask if they are also influencers in the draft co-authorship graph.

Thus, we create a draft co-authorship graph where each author is a node,
and draft co-authorship is an edge. We then measure influence with
betweenness centrality of the authors in each time period.
Table~\ref{tbl:correlation_sprmn} compares the top 20\% of influencers of
the mailing lists with those of the co-authorship graph.  We find a
significant overlap between both groups, ranging from 48.2\% to 67.2\%.
There is a significant ($p < 0.05$) positive correlation between the
rankings of the overlapping members of each community: participants that
are influential in the mailing lists are also likely to be influential in
draft authorship.

@@

We also look at the top 20\% participants from the co-authorship network
who are \emph{not} part of the top 20\% influencers in the email network in
Figure \ref{fig:non_overlapping_coauthornetwork_authors_emailnetwork}.
We observe that not all the prolific authors are that engaged in the email
discussion: 40\%-50\% of non-overlapping authors are ranked between 20th to
40th percentile of influence in the email network. These non-overlapping
authors are typically more junior with respect to their participation
duration.

@@

%..................................................................................................
\pb{Leadership roles} 
WG chairs are selected from the community, and we might expect those
selected to be influential in the community.
Figure~\ref{fig:wg_chair_before_after} shows that this is indeed the case:
87.5\% of WG chairs are in the top 20\% of mailing list influencers and
67.4\% are in the top 20\% of document authors in the year before they
became chairs. This also shows the impact of taking up a leadership role:
influence in both the mailing list and authorship communities grows in the
year after participants become a WG chair for the first time.

@@

%..................................................................................................
\pb{Organisations}  While participants in the IETF contribute as
individuals, they are usually affiliated with an organisation.  To study
the potential influence of organisations, for each draft we obtain the
authors' affiliations from the Datatracker. If this is not available, we
use the domain name of the authors' email addresses, which we map to the
relevant organisation (e.g., \texttt{@cisco.com} maps to Cisco). For generic
email addresses, such as \texttt{@gmail.com}, we use the participant name
if no affiliation is available.

Figure \ref{fig:coauthor_affiliation_network2019_otheryears} shows the ten
most frequent affiliated organisations of the top 20\% authors in the
period 2015-2019, and the number of authors affiliated with these
organisations in  each  period.  While the dominance of Cisco is clear,
other organisations, such as Huawei, have gained a larger presence over
time.  In general, a small number of organisations employ a significant
fraction of the influential participants in the standards process. If we
look at the period 2015-2019, for example, out of the 915 authors ranked in
the top 20\% of most influential authors, 342 belong to the ten most
frequently affiliated organisations, and nearly 253 are from Cisco, Huawei,
Ericsson, or Juniper alone.

@@

Further inspection of the collaboration trends reveals that the volume of
collaborations between authors from Huawei and other organisations appears
to be unaffected by the recent addition of Huawei to the U.S. Entity List
\cite{bis:2019:entity} as the trends in Figure~\ref{fig:heatmap_2019_coauthordraft_most_frequent_org}
also hold for 2020-2021.\footnote{\tiny\url{https://mailarchive.ietf.org/arch/msg/ietf-announce/0ywjgSS4LlO0DaWDoLJLRHxJdUk/}}

>>>>

While the top participants are defined by the betweenness centrality metric
(\S\ref{subsec:measuring_influence}) as being influential within the IETF
mailing lists, it is unclear how this influence translates into document
authorship and leadership roles within the IETF.  We explore these aspects
of influence in the following.

%..................................................................................................
\pb{Document authorship:}
The output of the IETF is technical standards and other documents published
in the RFC series. RFCs are developed by working groups from a sequence of
Internet-drafts, with individuals acting as named authors. Having identified
the most influential participants in the mailing list community, we can
determine whether or not these same participants are also the most active
authors.

Figure~\ref{fig:drafts_initiation_top_percentile} shows the distribution of
the proportion of documents authored by mailing list participants, sorted
by their influence rank. We observe that:
(i) influential participants in the email graph, i.e., those ranked top in
betweenness centrality, tend to write more drafts than do participants with
lower centrality; and
(ii) the number of drafts per participant is increasing over time for all
participants.
For example, during 2010-2019, each participant in the top 10 percentile,
authored 0.125\% to 0.175\% of the total number of drafts in that period.
We find that 32\%-42\% of the total drafts, in different years during this
period, were authored by the top 10\% most influential mailing list
participants.

@@

%..................................................................................................
\pb{Document co-authorship and correlation with the email graph:}
Figure~\ref{fig:drafts_initiation_top_percentile} showed that influential
participants on the mailing lists tend to author more drafts than others.
We next ask if they are also influencers in the draft co-authorship graph.

To do this, we create a draft co-authorship graph where each author is a
node, and draft co-authorship is an edge. We then measure influence with
betweenness centrality of the authors in each time period.
Table~\ref{tbl:correlation_sprmn} compares the top 20\% of influencers of
the mailing lists with those of the co-authorship graph.  We find a
significant overlap between both groups, ranging from 48.2\% to 67.2\%.
There is a significant ($p < 0.05$) positive correlation between the
rankings of the overlapping members of each community: participants that
are influential in the mailing lists are also likely to be influential in
draft authorship.

@@

We also look at the top 20\% participants from the co-authorship network
who are \emph{not} part of the top 20\% influencers in the email network in
Figure \ref{fig:non_overlapping_coauthornetwork_authors_emailnetwork}.
We observe that that are some prolific authors are that \emph{not} engaged
in the email discussion: 40\%-50\% of non-overlapping authors are ranked
between 20th to 40th percentile of influence in the email network. These
non-overlapping authors are typically more junior with respect to their
participation duration.

@@

%..................................................................................................
\pb{Influencers and leadership roles:} 
Working group chairs are selected from the community, so we might expect
those selected to be influential in the community.
Figure~\ref{fig:wg_chair_before_after} shows that this is indeed the case:
87.5\% of working group chairs are in the top 20\% of mailing list
influencers and 67.4\% are in the top 20\% of document authors in the year
before they became chairs. This also shows the impact of taking up a
leadership role: influence in both the mailing list and authorship
communities grows in the year after participants become a working group
chair for the first time.

@@

%..................................................................................................
\pb{Influence of organisations:}
While participants in the IETF are expected to contribute as individuals,
they are usually affiliated with an organisation.  To study the potential
influence of those organisations, for each draft we obtain the authors'
affiliations from the Datatracker. If this is not available, we use the
domain name of the authors' email addresses, which we map to the relevant
organisation (e.g., \texttt{@cisco.com} maps to Cisco). For generic email
addresses, such as \texttt{@gmail.com}, we use the participant name if no
affiliation is available.

Figure \ref{fig:coauthor_affiliation_network2019_otheryears} shows the ten
most frequent affiliated organisations of the top 20\% of draft authors in
the period 2015-2019, and the number of authors affiliated with these
organisations in  each  period.  While the dominance of Cisco is clear,
other organisations, such as Huawei, have gained a larger presence over
time.  In general, a small number of organisations employ a significant
fraction of the influential participants in the standards process. If we
look at the period 2015-2019, for example, out of the 915 authors ranked in
the top 20\% of most influential authors, 342 belong to the ten most
frequently affiliated organisations, and nearly 253 are from Cisco, Huawei,
Ericsson, or Juniper alone.

@@

Further inspection of the collaboration trends reveals that the volume of
collaborations between authors from Huawei and other organisations appears
to be unaffected by the addition of Huawei to the U.S. Entity List
\cite{bis:2019:entity} as the trends in Figure 
\ref{fig:heatmap_2019_coauthordraft_most_frequent_org} also hold for 2020-2021.
